{\LARGE Model Selection for Scalar-on-Function Regression \\ with Applications to Near-Infrared Spectroscopy}

### 2022-02-03 Paper/main.tex
Adding tables in seperat .tex, Further interpretation of results

		where $Z \sim \mathcal{N}(0,1)$. In total, we created four combinations for the simulations, using the two true coefficient functions and the two sets of responses. These four combinations are then used with a different number of monomial basis functions $ \in \{1,2, \dots, 6\}$, cubic bspline basis-function $\{5,6,...,18\}$ and fourier functions $\{1,3,...,25\}$ to predict the generated responses using the basis expansion approach and the FPCR approach. For the evaluation, we used the prediction MSE calculated by 10 fold cross-validation.
		 To obtain valid out of sample properties for the FPCR, within each of the ten 10 fold cross-validation splits, we first calculate the Functional Principal Components of the trainig-set $\mathcal{T}$ for each curve. These scores are then used to estimate the scores of the holdout set $\mathcal{H}$, $\hat{\xi}_{i}^{m, \mathcal{H}}$  by the equation:

		 {\color{orange}I guess we don't use the scores of the training set to estimate scores of test set but eigenfunction of traningset..}
		 {\color{blue} We should mention how we derived the original principal components. (bspline basis and number of basis functions and harmonics.) As we chose a high number for both to capture all features of the data set. As we explain the approximations for the estimation, this question could come up, so we have to make it specific.}

@@

	\paragraph{Monomial Basis}
	Due to the high collinearity of these basis{(\color{orange} How about stressing the higher order? Are they already highly correlated at lower orders?)}, simulations were conducted up until the sixth monomial basis, excluding the first one since this just represents a constant. For reasons outlined in {\color{green} Link chapter}, they are suited for the smooth function $f_1$ and show a better performance than bsplines for this coefficient function, but show the weakest performance for the coefficient function $f_2$. For $f_1$, the simulation selects 5(3) and for $f_2$ 5(5) monomial basis functions for the high(low) signal-to-noise ratios. This weakness is especially pronounced in the setup $f_2,Y_1$; for the high noise setup, $f_2, Y_2$ this weakness is still visible, but less strongly pronounced.

	\paragraph{Bspline Basis}
	Simulations with bspline basis functions were possible from 5 to 18, since from 18 onwards the simulations were running into problems concerning {\color{green} what happend here exactly?}
The smooth function $f_1$ requires only 5(4) Bspline basis functions, for the high(low)-signal-to-noise ratio to obtain the best fit for the bspline basis, which performs the worst for $f_1$. For $f_2$, 11(6) bspline basis functions are needed for the high(low) signal-to-noise ratios. For $f_2$, the bspline basis functions outperform the monomial basis, but come second to the fourier basis.

@@

	\subsubsection{Functional Principal Component Regression}
	The model which is used for the FPCR is described in \hyperref[fpc_exp_transf]{\textbf{Estimation using Functional Principal Components}}, in which we transform the observed functions to perform regression of a scalar on a countable sequence of regressors, which is then tractable with typical multivariate regression theory.

	\paragraph{Monomial Basis}
	\paragraph{Bspline Basis}
	\paragraph{Fourier Basis}

@@

	\subsection{Literature}
	\begin{itemize}
		\item \cite{shonkwiler_explorations_2009}
		\item R-packages: fda, refund, mgcv
	\end{itemize}

	\newpage
	\section{Application}
		The application uses the insights from the previous sections to predict the octane ratings of the introduced gasoline dataset. Following the results from the simulation study,... 
		\begin{itemize}
			\item {\color{green} incorporate results of simulation study: number of basis, components, etc...}
			\item point out difficulties estimating sderror
			\item describe setup and results
\end{itemize}

@@

\newpage
	\subsection{Simulation Study Results}

	\begin{table}[htb]
			\centering
			\caption{Monomial Basis Expansion Regression}
				\begin{tabular}{lllllll}
					\cline{1-5}
					 \boldmath{$f_1, Y_1$}                 & \boldmath{$f_1, Y_2$}                  & \boldmath{$f_2, Y_1$}                    & \boldmath{$f_2, Y_2$}               & \textbf{n\_basis} &  \\ \cline{1-5}
7.177234352026647                        & 131.92716520680838                        & 0.8936725847175324                         & 2.5839516380614462                        & 2       &  \\
4.170391190343292                        & {\color[HTML]{FE0000} 129.64258934345224} & 0.8122408608804015                         & 2.511772278865247                         & 3       &  \\
3.912749103345826                        & 130.1170783435205                         & 0.3807927171910713                         & 2.090474762002978                         & 4       &  \\
{\color[HTML]{FE0000} 3.638501968909747} & 130.59817144950233                        & {\color[HTML]{FE0000} 0.09216751463887264} & {\color[HTML]{FE0000} 1.8134256416939782} & 5       &  \\
6.016442343989896                        & 201.15534584991525                        & 0.7620805051836907                         & 3.3937342332820926                        & 6       & 
\end{tabular}
\end{table}

\vspace{2cm}

	\begin{table}[htb]
			\centering
			\caption{BSpline Basis Expansion Regression}
				\begin{tabular}{lllllll}
					\cline{1-5}
					 \boldmath{$f_1, Y_1$}                 & \boldmath{$f_1, Y_2$}                  & \boldmath{$f_2, Y_1$}                    & \boldmath{$f_2, Y_2$}               & \textbf{n\_basis} &  \\ \cline{1-5}
3.9127490963955247                        & {\color[HTML]{FE0000} 130.1170783090359} & 0.3807927168700936                          & 2.090474761101881                         & 4       &  \\
{\color[HTML]{FE0000} 3.6430480624640396} & 130.61094959896604                       & 0.09511820431115069                         & 1.8164314387816713                        & 5       &  \\
3.6542611008674575                        & 131.35354539843397                       & 0.07775281510299467                         & {\color[HTML]{FE0000} 1.8091346529425483} & 6       &  \\
3.6770545937377634                        & 132.14205183557064                       & 0.07518488287219192                         & 1.8168036577708813                        & 7       &  \\
3.710740869697218                         & 133.3653658811635                        & 0.05810108378740701                         & 1.8164984632128844                        & 8       &  \\
3.7192146351418724                        & 133.6814928169343                        & 0.05640282340664383                         & 1.8193083539013821                        & 9       &  \\
3.742010447810594                         & 134.5121710907784                        & 0.0521785231194299                          & 1.8257576297312081                        & 10      &  \\
3.7643964092462974                        & 135.29727347111702                       & {\color[HTML]{FE0000} 0.051902252769905684} & 1.8360966075573522                        & 11      &  \\
3.80109308638683                          & 136.5758146586355                        & 0.052040315671138586                        & 1.8531479959118535                        & 12      &  \\
3.834360204971099                         & 137.78406681481994                       & 0.052706573106132304                        & 1.869969847032031                         & 13      &  \\
3.862166917165765                         & 138.73431326927349                       & 0.053068654178390816                        & 1.8826933364948355                        & 14      &  \\
3.8685638723557956                        & 138.97427074660192                       & 0.0525955164519272                          & 1.8848221385552892                        & 15      &  \\
3.88505750612787                          & 139.57418873526626                       & 0.05283448721173191                         & 1.8933474257957223                        & 16      &  \\
3.912671952971778                         & 140.51490301138338                       & 0.053220205231306356                        & 1.9061942087104604                        & 17      &  \\
3.948125713522542                         & 141.80885020010606                       & 0.053584434348457485                        & 1.9231184762987383                        & 18      & 
\end{tabular}
\end{table}

\vspace{2cm}

	\begin{table}[htb]
			\centering
			\caption{Fourier Basis Expansion Regression}
				\begin{tabular}{lllllll}
					\cline{1-5}
					 \boldmath{$f_1, Y_1$}                 & \boldmath{$f_1, Y_2$}                  & \boldmath{$f_2, Y_1$}                    & \boldmath{$f_2, Y_2$}               & \textbf{n\_basis} &  \\ \cline{1-5}
3.6975216821108177                        & {\color[HTML]{FE0000} 129.19133735032568} & 0.6952364931614831                         & 2.394399533437222                         & 3       \\
{\color[HTML]{FE0000} 3.6347035142043147} & 130.59281554182246                        & 0.07418129892625985                        & 1.795816800944426                         & 5       \\
3.6762259845089287                        & 132.0824820138016                         & 0.05147130962678385                        & {\color[HTML]{FE0000} 1.7934269021325526} & 7       \\
3.718852537544315                         & 133.67575428625497                        & {\color[HTML]{FE0000} 0.05104970622270979} & 1.812908100982769                         & 9       \\
3.764506721475903                         & 135.26218648523314                        & 0.05146451789575423                        & 1.8346302450265894                        & 11      \\
3.810949927900835                         & 136.90281800923086                        & 0.05197380387819929                        & 1.8572425102431276                        & 13      \\
3.8567441840723165                        & 138.57257529224665                        & 0.05252120988973837                        & 1.8802107534919845                        & 15      \\
3.9061852571557436                        & 140.29178124765565                        & 0.05304327165297495                        & 1.9028253007284568                        & 17      \\
3.9551703628111876                        & 142.057266882177                          & 0.053647142499864804                       & 1.9271845824346567                        & 19     
\end{tabular}
\end{table}

@@

	\section{Proofs}
	\subsection{Lemma} \label{Proof1}
	The curves $X(t) \in \mathbb{L}^2[0,1]$ is expanded by the eigenfunctions $\{\nu_j\}$ as (\ref{KarhunenLoeve}). The coefficients $\xi_{j}$ of basis functions satisfy the following properties:
	\begin{enumerate}
		\item $\mathbb{E}\left[\xi^m(\omega)\right] = 0$
		\item $Cov\left(\xi^m(\omega), \xi^n(\omega)\right) = \delta^{m,n}\lambda^m$% if $m \neq n$
		\item $Var\left(\xi^m(\omega)\right) = \lambda^m$
	\end{enumerate}
	Remind that $\delta^{m,n} = 0$ if $m \neq n$, otherwise 1.
	\begin{proof}
		Assume that $F(t)$ is the centered process of $X(t)$, namely, $F(t) = X(t) - \int_{\Omega}X(t)dP(\omega)$. To obtain the first result, we can show that
		\begin{equation}\label{Lemma1}
			\begin{split}
				\mathbb{E}[\xi_{j}] = & \mathbb{E} \biggl\lbrack \int_{0}^{1} F(t) \nu_{j}(t)dt\biggr\rbrack\\
				= & \int_{\Omega} \int_{0}^{1} F(t) \nu_{j}(t) dt dP(\omega)\\
				= & \int_{0}^{1} \int_{\Omega} F(t) \nu_{j}(t) dP(\omega) dt \quad \text{(Fubini)}\\
				= & \int_{0}^{1} {\color{red}\int_{\Omega} F(t) dP(\omega)} \nu_{j}(t) dt\\
				= & \int_{0}^{1} {\color{red}\mathbb{E}[F(t)]} \nu_{j}(t) dt = 0
			\end{split}
		\end{equation}
		where $\mathbb{E}[F(t)]$ is 0 since $F(t)$ is a centered process.
		The second claim is proved as:
		\begin{equation}\label{Lemma2}
			\begin{split}
				\mathbb{E} [\xi_{j} \xi_{k}] = & \mathbb{E}  \biggl\lbrack \int_{0}^{1} F(s) \nu_{j}(s)ds \int_{0}^{1} F(t) \nu_{k}(t)dt  \biggr\rbrack\\
				= & \mathbb{E} \biggl\lbrack {\int_{0}^{1} \int_{0}^{1} F(s) \nu_{j}(s) F(t) \nu_{k}(t) ds dt} \biggr\rbrack \quad \text{(Fubini)}\\
				= & \int_{0}^{1} {\color{red}\int_{0}^{1} \mathbb{E}[{F(s)F(t)}] \nu_{j}(s)} \nu_{k}(t) {\color{red}ds} dt\\
				= & \int_{0}^{1} {\color{red}\left(\int_{0}^{1}c(s,t)\nu_{j}(s)ds \right)} \nu_{k}(t) dt \\
				= & \int_{0}^{1}{\color{red}[K\nu_{j}](t)}\nu_{k}(t)dt\\
				= & \langle K \nu_{j}, \nu_{k} \rangle\\
				= & \langle \lambda_{j} \nu_{j}, \nu_{k} \rangle = \delta_{j,k}\lambda_{j}
			\end{split}
		\end{equation}
		where $\delta_{j,k} = 1$ if $j = k$, otherwise 0. The result is produced from orthonormaility of the eigenfunctions. The last assertion is confirmed from the above two properties.
		\begin{equation}\label{Lemma3}
			Var[\xi_{i}] = \mathbb{E}\left[(\xi_{i} - \mathbb{E}[\xi_{i}])^{2}\right] = \mathbb{E}[\xi_{i}^{2}] = \lambda_{i}
		\end{equation}
		The original process also has the same properties as the centered one since
		\begin{equation}
			X(t) = F(t) + \mathbb{E}[X(t)] = \mu(t) + \sum_{j=1}^{\infty}\xi_{j}\nu_{j}(t)
		\end{equation}
	\end{proof}

>>>>

		where $Z \sim \mathcal{N}(0,1)$. In total, we created four combinations for the simulations, using the two true coefficient functions and the two sets of responses. These four combinations are then used with a different number of monomial basis functions $ \in \{1,2, \dots, 6\}$, cubic bspline basis-function $\{5,6,...,18\}$ and fourier functions $\{1,3,...,25\}$ to predict the generated responses using the basis expansion approach and the FPCR approach. For the evaluation, we used the prediction RMSE calculated by 10 fold cross-validation.
		 To obtain valid out of sample properties for the FPCR, within each of the ten 10 fold cross-validation splits, we first calculate the Functional Principal Components of the trainig-set $\mathcal{T}$ for each curve. These scores are then used to estimate the scores of the holdout set $\mathcal{H}$, $\hat{\xi}_{i}^{m, \mathcal{H}}$  by the equation:

		 {\color{blue} We should mention how we derived the original principal components. (bspline basis and number of basis functions and harmonics.) As we chose a high number for both to capture all features of the data set. As we explain the approximations for the estimation, this question could come up, so we have to make it specific.}

@@

	\paragraph{Monomial Basis}
	Due to the high collinearity of these basis, simulations were conducted up until the sixth monomial basis, excluding the first one since this just represents a constant. As outlined in {\color{green} Link chapter}, they seem suited for $f_1$ and show a better performance in the simulations than bsplines for this coefficient function, but show the weakest performance for the coefficient function $f_2$. For $f_1$, the simulation selects 5(3) and for $f_2$ 5(5) monomial basis functions for the high(low) signal-to-noise ratios. The outlined weakness due to collinearity is especially pronounced in $f_2,Y_1$; for the high noise setup, $f_2, Y_2$ this weakness is still visible, but less strongly pronounced.

	\paragraph{Bspline Basis}
	Simulations with bspline basis functions were possible from 5 to 18, since from 18 onwards the simulations were running into problems concerning {\color{green} what happend here exactly?}
The smooth function $f_1$ requires only 5(4) Bspline basis functions, for the high(low) signal-to-noise ratio to obtain the best fit for the bspline basis, which performs the worst for $f_1$. For $f_2$, 11(6) bspline basis functions are needed for the high(low) signal-to-noise ratios. For $f_2$, the bspline basis functions outperform the monomial basis, but come second to the fourier basis.

@@

	\subsubsection{Functional Principal Component Regression}
	The model which is used for the FPCR is described in \hyperref[fpc_exp_transf]{\textbf{Estimation using Functional Principal Components}}. Additionally to the selection of the truncation parameter $K$, the choice of the number of FPC´s adds to the complexity of the model since the eigenvalues and -functions from the decomposition are influenced by the choice of $K$ that was used to approximate the function.

	\paragraph{Monomial Basis}
	TBD
	\paragraph{Bspline Basis}
	For $f_1,Y_1$, the CV suggests a better model the more FPC´s are used. While the number of basis functions stays at 5 for two and three used FPC, it increases to 6 basis functions for $nharm$ = 4. In $f_1,Y_1$, the cross-validated MSE is only slightly affected by the number of FPC´s, but lowest for $nharm$ = 3, indicating the possibility of over-fitting for four FPC´s. $f_2,Y_1$ shows the strongest decrease of MSE in increasing FPC´s of all FPCR models fitted with BSplines (0.69614, 0.42504, 0.09497 for 2,3,4 FPC´s). For $f_2,Y_2$ the same holds true as in the setup with $f_2$ and the high signal-to-noise ration, but less pronounced. Similar to the two setups with $f_1$, the cross-validated number of basis functions is increasing in the number of FPC´s (4,6,8 for 2,3,4 FPC´s)\\
	The results overall might indicate that a higher number of FPC´s has a strong effect in the setup with high signal-to-noise ratio, while not reducing as much error in the setting with the higher noise.

	\paragraph{Fourier Basis}
	We observe similar behaviour as for the FPCR using BSplines. In $f_1,Y_1$, the MSE is deceasing in the number of FPC´s while the results in $f_1,Y_2$ might again indicate over fitting when using 4 FPC´s. Since it was not possible to calculate the cross-validated MSE for 3 basis functions for 4 FPC´s, the increasing number of basis functions as seen in the FPCR with BSplines can not be observed here. $f_2,Y_1$ displays again the greatest decline of MSE in the increasing number of FPC´s. The same behaviour can be seen in $f_2,Y_2$, but less pronounced. Both configurations of $f_2$ a re using 5,15,7 basis functions for 2,3,4 FPC´s: This behaviour is opposing to the previous reported increasing nature of basis functions in the number of FPC`s.

@@

	%\subsection{Literature}
	%\begin{itemize}
	%	\item \cite{shonkwiler_explorations_2009}
	%	\item R-packages: fda, refund, mgcv
	%\end{itemize}

	\newpage
	\section{Application}
		The application uses the methods from the previous sections to predict the octane ratings of the introduced gasoline dataset. The simulation study will contribute towards the reading of the results. 
		\begin{itemize}
			\item {\color{green} incorporate results of simulation study: 					number of basis, components, etc...}
			\item point out difficulties estimating sderror
			\item describe setup and results
		\end{itemize}

@@

\newpage
\include{tables.tex}

@@

	\section{Proofs}
	\subsection{Lemma} \label{Proof1}
	The curves $X(t) \in \mathbb{L}^2[0,1]$ is expanded by the eigenfunctions $\{\nu_j\}$ as (\ref{KarhunenLoeve}). The coefficients $\xi_{j}$ of basis functions satisfy the following properties:
	\begin{enumerate}
		\item $\mathbb{E}\left[\xi^m(\omega)\right] = 0$
		\item $Cov\left(\xi^m(\omega), \xi^n(\omega)\right) = \delta^{m,n}\lambda^m$% if $m \neq n$
		\item $Var\left(\xi^m(\omega)\right) = \lambda^m$
	\end{enumerate}
	Remind that $\delta^{m,n} = 0$ if $m \neq n$, otherwise 1.
	\begin{proof}
		Assume that $F(t)$ is the centered process of $X(t)$, namely, $F(t) = X(t) - \int_{\Omega}X(t)dP(\omega)$. To obtain the first result, we can show that
		\begin{equation}\label{Lemma1}
			\begin{split}
				\mathbb{E}[\xi_{j}] = & \mathbb{E} \biggl\lbrack \int_{0}^{1} F(t) \nu_{j}(t)dt\biggr\rbrack\\
				= & \int_{\Omega} \int_{0}^{1} F(t) \nu_{j}(t) dt dP(\omega)\\
				= & \int_{0}^{1} \int_{\Omega} F(t) \nu_{j}(t) dP(\omega) dt \quad \text{(Fubini)}\\
				= & \int_{0}^{1} {\color{red}\int_{\Omega} F(t) dP(\omega)} \nu_{j}(t) dt\\
				= & \int_{0}^{1} {\color{red}\mathbb{E}[F(t)]} \nu_{j}(t) dt = 0
			\end{split}
		\end{equation}
		where $\mathbb{E}[F(t)]$ is 0 since $F(t)$ is a centered process.
		The second claim is proved as:
		\begin{equation}\label{Lemma2}
			\begin{split}
				\mathbb{E} [\xi_{j} \xi_{k}] = & \mathbb{E}  \biggl\lbrack \int_{0}^{1} F(s) \nu_{j}(s)ds \int_{0}^{1} F(t) \nu_{k}(t)dt  \biggr\rbrack\\
				= & \mathbb{E} \biggl\lbrack {\int_{0}^{1} \int_{0}^{1} F(s) \nu_{j}(s) F(t) \nu_{k}(t) ds dt} \biggr\rbrack \quad \text{(Fubini)}\\
				= & \int_{0}^{1} \int_{0}^{1} {\color{red}\mathbb{E}[{F(s)F(t)}]} \nu_{j}(s) \nu_{k}(t) ds dt\\
				= & \int_{0}^{1} {\color{red}\left(\int_{0}^{1}c(s,t)\nu_{j}(s)ds \right)} \nu_{k}(t) dt \\
				= & \int_{0}^{1}{\color{red}[K\nu_{j}](t)}\nu_{k}(t)dt\\
				= & \langle K \nu_{j}, \nu_{k} \rangle\\
				= & \langle \lambda_{j} \nu_{j}, \nu_{k} \rangle = \lambda_{j} \quad \text{if $j = k$, otherwise 0}
			\end{split}
		\end{equation}
		where the result is produced from orthonormaility of the eigenfunctions. The last assertion is confirmed from the above two properties.
		\begin{equation}\label{Lemma3}
			Var[\xi_{i}] = \mathbb{E}\left[(\xi_{i} - \mathbb{E}[\xi_{i}])^{2}\right] = \mathbb{E}[\xi_{i}^{2}] = \lambda_{i}
		\end{equation}
		The original process also has the same properties as the centered one since
		\begin{equation}
			X(t) = F(t) + \mathbb{E}[X(t)] = \mu(t) + \sum_{j=1}^{\infty}\xi_{j}\nu_{j}(t)
		\end{equation}
	\end{proof}
